---
output: html_document
---
#Analyzing Portuguese "Vinho Verde" Red Wine Quality
###by Bhavin V. Choksi

***

This project analyzes the physicochemical properties that affect the quality of
1599 variants of the Portuguese "Vinho Verde" red wine.

The physicochemical properties in the data set are based on objective tests.

The quality of each wine is graded from 0 (very bad) to 10 (very excellent),
based on the median of at least 3 evaluations by wine experts.

My objective is to determine which of the physicochemical properties affect wine quality, and then build a linear model based on those factors to predict quality.

Property|Unit|Description
--|--|--
Fixed acidity|gm/L|Most acids involved with wine or fixed or nonvolatile (do not evaporate readily).
Volatile acidity|gm/L|The amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste.
Citric acid|gm/L|Found in small quantities, citric acid can add 'freshness' and flavor to wines.
Residual sugar|gm/L|The amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet.
Chlorides|gm/L|The amount of salt in the wine.
Free sulfur dioxide|mg/L|The free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine.
Total sulfur dioxide|mg/L|Amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine.
Density|gm/mL|The density of wine is close to that of water depending on the percent alcohol and sugar content.
pH||Describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale.
Sulphates|gm/L|A wine additive which can contribute to sulfur dioxide gas (S02) levels, which acts as an antimicrobial and antioxidant.
Alcohol|%|The percent alcohol content of the wine. 

[Wine Quality Data Set Information](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt)

***

###Data

```{r echo = FALSE, warning = FALSE}
#Loading data
wqr <- read.csv(file = "~/Documents/Personal/Data Analyst Nanodegree/P3 Data Analysis with R/Project/wineQualityReds.csv")
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
```

Confirming that all 1599 rows were loaded.
```{r echo = FALSE, warning = FALSE}
nrow(wqr)
```

Confirming that all columns were loaded.
```{r echo = FALSE, warning = FALSE}
colnames(wqr)
```

Sample data shows that quality has discrete numeric values, and all
physicochemical properties have continuous numeric values.

```{r echo = FALSE, warning = FALSE}
head(wqr)
```

```{r echo = FALSE, warning = FALSE}
#Removing the first "serial no." column
wqr$X <- NULL

#Categorizing wine qualities as Low, Medium and High
wqr$quality.category[wqr$quality == 3 | wqr$quality == 4] <- "Low"
wqr$quality.category[wqr$quality == 5 | wqr$quality == 6] <- "Medium"
wqr$quality.category[wqr$quality == 7 | wqr$quality == 8] <- "High"
wqr$quality.category <- factor(wqr$quality.category,
                               levels = c("Low", "Medium", "High"))
```

***

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
```

##Univariate Plots and Analysis

###Wine Quality

A summary and plot of quality shows that wines in the data set have grades
between 3 and 8. None of the wines are close to being very bad or very
excellent.

For the purpose of analysis, I have categorized grades 3 and 4 as Low, grades 5
and 6 as Medium, and grades 7 and 8 as High.

About 4% of the wines are of Low quality.
82.5% of the wines are of Medium quality.
13.5% of the wines are of High quality.

```{r echo = FALSE, warning = FALSE}
#Quality
table(wqr$quality)
summary(wqr$quality)

#Quality Category
table(wqr$quality.category)
prop.table(table(wqr$quality.category))
```

```{r echo = FALSE, warning = FALSE}
ggplot(aes(x = factor(quality)), data = wqr) +
  geom_histogram() +
  xlab("quality")
```

Some wine qualities are outliers (grades 3 and 8) for the provided data.
However, they do belong in the data set for further analysis as each wine
undergoes at least three evalauations, and hence cannot be errors.

```{r echo = FALSE, warning = FALSE}
#Wine quality outliers
lower.bound <- quantile(wqr$quality, 0.25) - IQR(wqr$quality)
upper.bound <- quantile(wqr$quality, 0.75) + IQR(wqr$quality)

sort(unique(wqr$quality[wqr$quality < lower.bound | wqr$quality > upper.bound]))
```

###Fixed Acidity

The values look normally distributed with few outliers, with most wines here
having fixed acidity in the range of 5 to 11 gm/L.

```{r echo = FALSE, warning = FALSE}
summary(wqr$fixed.acidity)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = fixed.acidity, data = wqr, binwidth = .1) +
  scale_x_continuous(breaks = seq(4, 16, 1))
```

###Volatile Acidity

Volatile acidity also seems to be normally distributed, most wines here having
acetic acid in the range of 0.2 to 0.9 gm/L.

The mean and median are close, 0.53 and 0.52 respectively.

```{r echo = FALSE, warning = FALSE}
summary(wqr$volatile.acidity)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = volatile.acidity, data = wqr, binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 1.6, .1))
```

###Citric Acid

Over 8% of the wines here (132 of 1599) have no citric acid, the rest having
less than 0.8 gm/L.

Mean and median are close, 0.27 and 0.26 respectively.

```{r echo = FALSE, warning = FALSE}
table(wqr$citric.acid == 0)
summary(wqr$citric.acid)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = citric.acid, data = wqr, binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 1, 0.1))
```

###Residual Sugar

Most wines here have between 1 and 3 gm/L of residual sugar, with some outliers
having up to 15 gm/L.

```{r echo = FALSE, warning = FALSE}
summary(wqr$residual.sugar)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = residual.sugar, data = wqr, binwidth = .1) +
  scale_x_continuous(breaks = seq(0, 16, 1))
```

###Chlorides

Most wines here have 0.04 to 0.12 gm/L of salt, with a few outliers.

```{r echo = FALSE, warning = FALSE}
summary(wqr$chlorides)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = chlorides, data = wqr, binwidth = .001) +
  scale_x_continuous(breaks = seq(0, 0.6, .05))
```

###Free and Total Sulfur Dioxide

The distribution of values is long tailed, with most wines here having free
sulfur dioxide in the range of 3 to 40 mg/L, and total sulfur dioxide in the
range of 6 to 150 mg/L.

Only wines with over 50 ppm (mg/L) of free sulfur dioxide concentrations are
detectable to affect the nose and taste. Just 1% (16 of 1599) of the wines here
have free sulfur dixoide concentrations above 50 ppm.

```{r echo = FALSE, warning = FALSE}
summary(wqr$free.sulfur.dioxide)
table(wqr$free.sulfur.dioxide > 50)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = free.sulfur.dioxide, data = wqr, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 75, 5))
```

```{r echo = FALSE, warning = FALSE}
summary(wqr$total.sulfur.dioxide)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = total.sulfur.dioxide, data = wqr, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 300, 20))
```

###Density

Wine density looks normally distributed, in a close range between
0.990 and 1.004 gm/mL.

```{r echo = FALSE, warning = FALSE}
summary(wqr$density)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = density, data = wqr, binwidth = .001) +
  scale_x_continuous(breaks = seq(0.990, 1.004, .001))
```

###pH

pH levels look normally distributed, mainly in the range of 3.0 to 3.6.

The mean and median are identical at 3.31.

```{r echo = FALSE, warning = FALSE}
summary(wqr$pH)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = pH, data = wqr, binwidth = .01) +
  scale_x_continuous(breaks = seq(2, 4, .1))
```

###Sulphates

Sulphates content is mostly between 0.45 to 0.9 gm/L. There are some outliers on
the higher end of the value range.

The mean and median are close, 0.62 and 0.6581 respectively.

```{r echo = FALSE, warning = FALSE}
summary(wqr$sulphates)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = sulphates, data = wqr, binwidth = .01) +
  scale_x_continuous(breaks = seq(0, 2, .1))
```

###Alcohol

Most wines here have an alcohol content between 9 and 13%.

Mean and median are a little over 10%.

```{r echo = FALSE, warning = FALSE}
summary(wqr$alcohol)
```

```{r echo = FALSE, warning = FALSE}
qplot(x = alcohol, data = wqr, binwidth = .1) +
  scale_x_continuous(breaks = seq(8, 14, 1))
```

***

##Bivariate Plots and Analysis

###Analyzing Physicochemical Properties vs. Quality

The following charts box plot each property against quality. The pattern on the
charts indicate a linear relation between quality and some of the properties
such as:

- volatile acidity
- citric acid
- density
- sulphates
- alcohol

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Exploring linear relationship between quality and each property,
#with layers to aid with overplotting
ggplot(aes(factor(quality), fixed.acidity), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), volatile.acidity), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), citric.acid), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), residual.sugar), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), chlorides), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), free.sulfur.dioxide), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), total.sulfur.dioxide), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), density), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), pH), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), sulphates), data = wqr) +
  geom_boxplot()

ggplot(aes(factor(quality), alcohol), data = wqr) +
  geom_boxplot()
```

###Exploring Conventional Wisdom

###Citric Acid and Quality

The data set description states that "citric acid can add 'freshness' and flavor
to wines". The box plot does indicate that, but I want to explore it further.

About 8% of the wines here (132 of 1599) have no citric acid. I want to compare
the distribution of grades for wines with and without citric acid.

The charts below show that only 6% of wines with no citric acid (8 of 132) are
of High quality (grade 7 or better).

Over 14% of wines with citric acid (209 of 1467) are of High quality.

The data analysis seems to confirm that the presence of citric acid does
influence wine quality positively.

```{r echo = FALSE, message = FALSE, warning = FALSE}
wqr$has.citric.acid <- "Has citric acid"
wqr$has.citric.acid[wqr$citric.acid == 0] <- "No citric acid"
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(aes(factor(quality), ), data = wqr) +
  geom_histogram() +
  facet_wrap(~ has.citric.acid)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
table(wqr$quality[wqr$citric.acid > 0])
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
table(wqr$quality[wqr$citric.acid == 0])
```

###Free Sulfur Dioxide and Quality

The data set description states that "at free SO2 concentrations over 50 ppm,
SO2 becomes evident in the nose and taste of wine". Only 1% of the wines here
(16 of 1599) meet that criteria, but I want to explore if free SO2 could impact
wine quality.

About 13.5% of the wines here (215 of 1581) with free SO2 concentrations below
50 ppm are of High quality (grade 7 or better).

On the other hand, only 12.5% of the wines here (2 of 16) with free SO2
concentrations over 50 ppm are of High quality.

Although this finding contradicts conventional wisdom, it is worth noting that
no wine here with over 50 ppm free SO2 concentration is of Low quality (grade
4 or lower).

```{r echo = FALSE, message = FALSE, warning = FALSE}
wqr$free.sulfur.dioxide.level <- "Free SO2 less than 50 ppm"
wqr$free.sulfur.dioxide.level[wqr$free.sulfur.dioxide > 50] <-
  "Free SO2 more than 50 ppm"
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(aes(factor(quality), ), data = wqr) +
  geom_histogram() +
  facet_wrap(~ free.sulfur.dioxide.level)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
table(wqr$quality[wqr$free.sulfur.dioxide < 50])
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
table(wqr$quality[wqr$free.sulfur.dioxide > 50])
```

###Statistical Measures

The correlation coefficients confirm that volatile acidity, sulphates and alcohol  influence wine quality to an extent that could be significant, and citric acid
and density to a lesser degree.

```{r echo = FALSE, warning = FALSE}
#Initialize
property <- character(11)
r <- numeric(11)
min <- numeric(11)
max <- numeric(11)
col.names <- colnames(wqr)

#Create a tabular view to see estimated coefficients and min/max values
#for each property at a glance
for(i in 1:11) {
  property[i] <- col.names[i]
  r[i] <- as.numeric(cor.test(wqr$quality, wqr[[col.names[i]]])$estimate)
  
  summ <- summary(wqr[[col.names[i]]])
  min[i] <- summ[1]
  max[i] <- summ[6]
}

properties <- data.frame(property, r, min, max)
properties
```

###Analyzing Relationship Between Physicochemical Properties

###Citric Acid and Fixed Acidity

Citric acid being non-volatile has a positive linear relationship with fixed
acidity.

From the scales it seems citric acid is a small part of overall fixed acidity.

[Fixed acidity in wine.](http://waterhouse.ucdavis.edu/whats-in-wine/fixed-acidity)

```{r echo = FALSE, warning = FALSE}
#Acidity and citric acid
ggplot(aes(fixed.acidity, citric.acid), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(5, 11) +
  stat_smooth(method = "lm", colour = "red")
```

###Citric Acid and Volatile Acidity

Acetic acid, the volatile acid in wine, seems to be lower in wines when
citric acid is higher.

```{r echo = FALSE, warning = FALSE}
ggplot(aes(volatile.acidity, citric.acid), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(0.2, 0.9) +
  stat_smooth(method = "lm", colour = "red")
```

###Fixed Acidity and Density

There is a positive linear relationship between fixed acidity and density.

This is to be expected as the fixed acids found in wine are denser than water
(1 gm/mL).

Fixed Acid in Wine|Density
--|--
Tartaric Acid|1.79 gm/mL
Malic Acid|1.61 gm/mL
Citric Acid|1.67 gm/mL
Succinic Acid|1.56 gm/mL

```{r echo = FALSE, warning = FALSE}
ggplot(aes(fixed.acidity, density), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(5, 11) +
  stat_smooth(method = "lm", colour = "red")

ggplot(aes(citric.acid, density), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(0, 0.8) +
  stat_smooth(method = "lm", colour = "red")
```

###Acidity and pH

As expected, pH levels have an inverse linear relationship with fixed acidity.
The lower the pH level, the more acidic the solution.

```{r echo = FALSE, warning = FALSE}
ggplot(aes(fixed.acidity, pH), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(5, 11) +
  stat_smooth(method = "lm", colour = "red")

ggplot(aes(citric.acid, pH), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(0, 0.8) +
  stat_smooth(method = "lm", colour = "red")
```

###Residual Sugar and Chlorides, and Density

The higher the residual sugar and salt in a wine, the denser it seems to be.

Measuring wine density is a method for categorizing it as Dry, Medium-sweet
or Sweet.

[Wine fermentation.](http://www.eng.umd.edu/~nsw/ench485/lab12.htm)

```{r echo = FALSE, warning = FALSE}
ggplot(aes(residual.sugar, density), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(1, 3) +
  stat_smooth(method = "lm", colour = "red")

ggplot(aes(chlorides, density), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(0.04, 0.12) +
  stat_smooth(method = "lm", colour = "red")
```

###Density and Alcohol

The plot shows an inverse linear relationship between alcohol and density,
the higher the alcohol content the lower the density.

The fermentation process converts sugars in grape juice to ethanol (ethyl alcohol). Density of ethanol is 0.789 gm/mL, which is lower than that of water (1 gm/mL).

[Wine fermentation.](http://www.eng.umd.edu/~nsw/ench485/lab12.htm)

```{r echo = FALSE, warning = FALSE}
ggplot(aes(density, alcohol), data = wqr) +
  geom_point(alpha = 1/2, position = "jitter") +
  xlim(.993, 1.001) +
  stat_smooth(method = "lm", colour = "red")
```

***

##Multivariate Plots and Analysis

###Density Estimates by Quality

These density charts indicate that there is a higher probability that wines
with better grades have higher alcohol, sulphates and citric acid content but
lower volatile acidity.

```{r echo = FALSE, message = FALSE, warning = FALSE}
qplot(alcohol, data = wqr, geom = "density", color = as.factor(quality))

qplot(sulphates, data = wqr, geom = "density", color = as.factor(quality))

qplot(volatile.acidity, data = wqr, geom = "density", color = as.factor(quality))

qplot(citric.acid, data = wqr, geom = "density", color = as.factor(quality))

qplot(density, data = wqr, geom = "density", color = as.factor(quality)) +
  xlab("wine.density")
```

###Histograms by Quality

The three histograms below confirm that wines of lower quality tend to have
lower alcohol and sulphates content but higher volatile acidity, and vice versa.

```{r echo = FALSE, message = FALSE, warning = FALSE}
qplot(alcohol, data = wqr, fill = as.factor(quality))

qplot(sulphates, data = wqr, fill = as.factor(quality))

qplot(volatile.acidity, data = wqr, fill = as.factor(quality))
```

On the following histograms, wines of all qualities seem to be spread across the
range of values of citric acid and density, indicating a weaker correlation
between those properties and quality.

```{r echo = FALSE, message = FALSE, warning = FALSE}
qplot(citric.acid, data = wqr, fill = as.factor(quality))

qplot(density, data = wqr, fill = as.factor(quality))
```

***

##Final Plots and Summary

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(RColorBrewer)
```

Plots, statistical measures and analysis thus far seem to indicate that there
is a strong correlation between the three physicochemical properties alcohol,
volatile acidity and sulphates, and wine quality.

The following plots provide further confirmation that these three properties
can be significant in predicting wine quality.

This plot indicates that wines of High quality tend to have higher
alcohol content and lower volatile acidity.

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(aes(alcohol, volatile.acidity, colour = quality.category), data = wqr) +
  geom_jitter() +
  xlim(9, 13) +
  ylim(0.2, 0.9) +
  scale_colour_brewer(type = "div")
```

This scatter plot indicates that Low and Medium quality wines are concentrated
at points where sulphates content is lower and volatile acidity is higher.

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(aes(sulphates, volatile.acidity, colour = quality.category), data = wqr) +
  geom_jitter() +
  xlim(0.45, 0.9) +
  ylim(0.2, 0.9) +
  scale_colour_brewer(type = "div")
```

This chart shows that as alcohol and sulphates content increases, wine quality
gets better.

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(aes(alcohol, sulphates, colour = quality.category), data = wqr) +
  geom_jitter() +
  xlim(9, 13) +
  ylim(0.45, 0.9) +
  scale_colour_brewer(type = "div")
```

###Interpreting the Charts and Estimated Coefficients
Property|Linear Relation|Comments
--|--|--
Fixed acidity|Weak, positive.|-
Volatile acidity|Medium to strong, negative.|Acetic acid at high levels can lead to an unpleasant, vinegar taste.
Citric acid|Weak to medium, positive.|Citric acid can add 'freshness' and flavor to wines.
Residual sugar|Very weak, positive.|All wines in the data set are fairly dry.<br>The highest residual sugar level is 15.50 gm/L.<br>Only wines over 45 gm/L are considered sweet.
Chlorides|Weak, negative.|-
Free sulfur dioxide|Very weak, negative.|-
Total sulfur dioxide|Weak, negative.|-
Density|Weak, negative.|Density of water is 0.99997 gm/mL.<br>All wines in the data set are close to the density of water (0.99 to 1.004).
pH|Very weak, negative.|On a scale of 0 (very acidic) to 14 (very basic) most wines are between 3-4.<br>All wines in the data set are between 2.74 and 4.01.
Sulphates|Medium, positive.|Sulphates contribute to sulfur dioxide gas (S02) levels, which acts as an antimicrobial and antioxidant.<br>Interestingly, free and total sulfur dioxide levels do not seem to impact quality.
Alcohol|Strong, positive.|-

###Building a model

Let's build a linear model to predict wine quality using the following properties
that have a medium to strong linear relation with quality as predictor variables:

- alcohol
- volatile acidity
- sulphates

We need two distinct samples from red wine quality data. One sample will be used to train the linear model. The other sample will be used to test the model, and compare its results with the actual evaluation by wine experts.

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(memisc)
```

```{r}
#1,500 rows in training set
#99 rows in test set
set.seed = 1056

sample.indices = sample(1:nrow(wqr), 1500)

training <- wqr[sample.indices, ]
test <- wqr[-sample.indices, ]

#Linear model
m1 <- lm(quality ~ alcohol, data = training)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + sulphates)
```

The linear model seems to be a good fit for the data based on the summary below:

- The R^2 value indicates that 33% of wine quality is due to its three properties - alcohol, volatile acidity and sulphates.

- The R^2 value of a model with only alcohol as a predictor variable indicates that 22% of wine quality is due to alcohol alone.

- Three significance stars (***) next to each property indicate that it is unlikely that no relationship exists between them and wine quality.

- A p-value of 0.000 for each property indicates a very low probability that they are not relevant in predicting wine quality.

```{r}
mtable(m1, m2, m3)
```

###Testing the model

A test of the model results in residuals that are fairly normally distributed, again indicating that the three properties are significant in predicting wine quality.

```{r}
#Predict
estimate <- predict(m3, newdata = test, interval = "prediction", level = 0.95)
estimate <- data.frame(estimate)
estimate$actual.quality <- NA
estimate$actual.quality <- test$quality
estimate$residual <- NA
estimate$residual <- estimate$fit - estimate$actual.quality
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Plot residuals
ggplot(aes(x = residual), data = estimate) +
  geom_density(fill = "red")
```

***

##Reflection

I started out with an analysis of each individual data element to get an idea
of the nature and distribution of its values. Univariate analysis indicated
that most wines are of Medium quality, and none of the wines have extremely low
or high ratings. Certain properties, such as citric acid and free sulfur dioxide,
that can positively impact quality were not found to be prevalent at high
rates or levels in the wines.

The analysis then progressed to test the impact of each physicochemical property
on quality. Alcohol, volatile acidity, sulphates, citric acid and density were
found to have a linear relationship with quality.

Relationships between each individual property were also analysed, revealing
correlation between density and alcohol, density and residual sugars, fixed
acidity and citric acid, besides others. Many of those relationships could be
explained by their physical and chemical attributes.

Multivariate analysis on alcohol, volatile acidity, sulphates, citric acid and
density revealed that citric acid and density did not have as strong a linear
relationship on wine quality as the other three properties. The correlation
coefficients confirmed that finding.

A linear model was built using alcohol, volatile acidity and sulphates as
predictor variables. The table of estimates and a plot of the residuals
indicated that the model was a good fit.

However, only 33% of wine quality is due to those three properties. It seems
natural that more than just 3 of 11 physicochemical properties of wine should
determine quality. A larger data set with a greater range of values for certain properties such as citric acid and free sulfur dioxide may allow us to use more predictor variables and build a better model.